---
title: "FItness Device Analysis"
author: "Adithyaa"
date: "20/10/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
# Overview

In this project, the aim is to analyse the data and classify it into A,B,C,D or E.


# Importing the necessary libraries

```{r}
library(readr) ##used for reading csv files in R
library(caret) ##classification and regression trees
library(rpart)
library(randomForest)
```


# Obtaining the data

## Reading the training dataset

```{r}
train <- read_csv("C:/Users/adith/Documents/PML_Project/train.csv")
summary(train)
head(train)
```



## Reading the test dataset

```{r}
test <- read_csv("C:/Users/adith/Documents/PML_Project/test.csv")
summary(test)
head(test)
```


# Performing Exploratory Data Analysis

```{r}
str(train) ## to obtain the type of each column
dim(train) ## checking the number of rows and columns
```

The training dataset contains 19622 rows and 160 columns

```{r}
table(train$classe) ##taking count for final results
```

# Data Cleaning

## Dealing with missing values

```{r}
na_count = sapply(1:dim(train)[2],function(x)sum(is.na(train[,x])))
na_list = which(na_count > 0)
colnames(train[,c(1:7)])
```

Now we know how many missing values we have and where they are. The next step is to remove these from the dataset.

## Removing the null values from both training and test sets

```{r}
train = train[,-na_list]
train = train[,-c(1:7)]
train$classe = factor(train$classe)
test = test[,-na_list]
test = test[,-c(1:7)]
```

# Model Building

## Decision Tree

```{r}
dtClassifier = rpart(formula = classe~.,data=train)
dt_y_prediction = predict(dtClassifier, train)
print(dt_y_prediction)

dt_y_prediction_test = predict(dtClassifier, test, type = "class")
print(dt_y_prediction_test)
```

## Random Forest

```{r}
rfClassifier = randomForest(formula = classe~.,data=train, ntree = 10)
rf_y_prediction = predict(rfClassifier, train)
print(rf_y_prediction)

rf_y_prediction_test = predict(rfClassifier, test)
print(rf_y_prediction_test)
```




